---
title: "Title of Your Report"
author: "Boyuan Cui, Zhaocheng Li"
date: "Oct 19, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(brms)
library(knitr)
```


## Abstract

(Here is where you give a brief (one paragraph overview of your entire paper). This should include some background/introduction, some methodology, results and conclusions.)

The report explores the impact of marital status, education level, individual income, self rated health, and average working hours on the well-being of individuals based on the 2017 General Social Survey data of Canada. The report analyzes the correlation through Bayesian linear model and uses Gibbs sampling to estimate the parameters of the model. We discovered that health, income, and certain marital status (married, living common law, and divorced) have a positive influence on the well-being, while education has a negative impact. Working hours and other marital status (widowed and separated) has no significant influence on the well-being.

## Introduction

(Here is where you should give insight into the setting and introduce the goal of the analysis. Here you can introduce ideas and basic concepts regarding the study setting and the potential model. Again, this is the introduction, so you should be explaining the importance of the work that is ahead and hopefully build some suspense for the reader. You can also highlight what will be included in the subsequent sections.)

Starting from 1985, Statistics Canada conducted General Social Survey (GSS) every year to collect cross-sectional data on non-institutionalized Canadian citizens over 15 years old to address and discover new  social issues. One of the primary objective of GSS is to monitor factors that affect Canadians' well-beings of that specific year (reference to user guide). We believe that an important indicator of a person's well-being is his/her own attitude towards life. If we are aware of what aspects promote or damage one's feeling of life, the government and any social agencies will have better insights on what they should do to make the society better off. We believe it is important to know which factors are influential on people's feeling of life as a whole, and we may find some unexpected discoveries.

The following sections will specifically explore the impact of marital status, education level, individual income, self rated health, and average working hours on one's feeling of life, and we will apply a Bayesian linear model to analyze the relationship.


## Data

(Introduce the data, explain why it was selected. Make sure to comment on important features and highlight any potential drawbacks to the data.)
```{r, echo=FALSE, include=FALSE}
gss <- read_csv("gss.csv")
gss_data <-  select(gss, feelings_life, marital_status, education, average_hours_worked, self_rated_health, income_respondent) %>% na.omit()


health_numeric <- case_when(gss_data$self_rated_health=="Excellent" ~ 5,
          gss_data$self_rated_health=="Very good" ~ 4,
          gss_data$self_rated_health=="Good" ~ 3,
          gss_data$self_rated_health=="Fair" ~ 2,
          gss_data$self_rated_health=="Poor" ~ 1
          )
          

work_hours_numeric = case_when(gss_data$ average_hours_worked=="50.1 hours and more" ~ 5,
                               gss_data$ average_hours_worked=="40.1 to 50.0 hours" ~ 4,
                               gss_data$ average_hours_worked=="30.0 to 40.0 hours" ~ 3,
                               gss_data$ average_hours_worked=="0.1 to 29.9 hours" ~ 2,
                               gss_data$ average_hours_worked=="0 hour" ~ 1)


                              
IncomeRes_numeric = case_when(gss_data$income_respondent=="$125,000 and more" ~ 6,
          gss_data$income_respondent=="$100,000 to $ 124,999" ~ 5,
          gss_data$income_respondent=="$75,000 to $99,999" ~ 4,
          gss_data$income_respondent=="$50,000 to $74,999" ~ 3,
          gss_data$income_respondent=="$25,000 to $49,999" ~ 2,
          gss_data$income_respondent=="Less than $25,000" ~ 1
          )
edu = case_when(gss_data$education=="University certificate, diploma or degree above the bach..." ~ 7,
          gss_data$education=="Bachelor's degree (e.g. B.A., B.Sc., LL.B.)" ~ 6,
          gss_data$education=="University certificate or diploma below the bachelor's level" ~ 5,
          gss_data$education=="College, CEGEP or other non-university certificate or di..." ~ 4,
          gss_data$education=="Trade certificate or diploma" ~ 3,
          gss_data$education=="High school diploma or a high school equivalency certificate" ~ 2,
          gss_data$education=="Less than high school diploma or its equivalent" ~ 1
          )

gss_numeric = data.frame(feelings_life = gss_data$feelings_life, marriage = factor(gss_data$marital_status), education=edu, health=health_numeric, income_respondent=IncomeRes_numeric,
                         work_hours=work_hours_numeric) %>% na.omit()
gss_numeric = within(gss_numeric, marriage <- relevel(marriage, ref = 5)) # make Single marital status as reference

```


### Data Extraction

We first obtain the GSS data file from the CHASS website ('AAIpsmGG.csv'), and clean the data with code from 'gss_cleaning.R' (written by professor Rohan Alexander and Sam Caetano) to make the raw data more interpretable. The data file after cleaning is 'gss.csv'

Then, we select our interest variables--marital status, education level, individual income, family income, self rated health, average working hours, and one's feeling of life--from the cleaned data, and we remove every case with NA or "Don't know" responses from the dataset. For the convenience of modeling, we will convert some categorical data to numeric variables. We will briefly explain each variable in the following section.

Below is an overview of the final dataset:
```{r, echo=FALSE}
head(gss_numeric) %>% tibble() %>% kable(align = 'lccrr', caption='Glimpse of Dataset')
```

### Variable Explanations

__Feeling of life__ (code in raw data: SLM_01, name in 'gss.csv': feelings_life) is a continuous numeric data ranging from 0 to 10 that indicates the respondent's feeling of life as a whole. 0 means very dissatisfied and 10 means very satisfied (reference from data).

__Marital status__ (code in raw data: MARSTAT, name in 'gss.csv': marital_status) is a categorical data indicating whether the respondent is divorced, living common-law, married, separated, single, or widowed [1].

We will not convert this data to numeric, since it is not intuitive to differentiate marital status into different levels. It is not reasonable to say that married is better than single, or vice versa, so we will leave this variable as categorical. But for concision purpose, we will give each marital status a code number for future reference:

```{r, echo=FALSE}
tibble('Marital status'=levels(gss_numeric$marriage),'Reference Code'=c(1:6)) %>% kable(align = 'lccrr', caption='Marital Status Reference')
```

__Education level__ (code in raw data: EHG3_01B, name in 'gss.csv': education) is a categorical data that indicates the highest education the respondent acquires [1]. Because the question asks for the highest education level achieved, it is reasonable to use numeric data to represent different levels of education. The variable after transformation will range from 1-7, representing from the lowest to the highest education level (we use the order of education level from the original GSS). Below is a reference table:
```{r, echo=FALSE}
edu_name = c("Less than high school diploma or its equivalent","High school diploma or a high school equivalency certificate","Trade certificate or diploma","College, CEGEP or other non-university certificate or diploma (other than trades certificates or diplomas)","University certificate or diploma below the bachelor's level","Bachelor's degree (e.g. B.A., B.Sc., LL.B.)","University certificate, diploma or degree above the bachelor’s level")
rank = group_by(gss_numeric, education) %>% summarise(.groups = 'drop')
rank = rank$education
tibble(`Education Level`=edu_name, `Numeric Rank`=rank) %>% kable(align = 'lccrr', caption='Education Level Reference')
```


__Individual income__ (code in raw data: TTLINCG2, name in 'gss.csv': income_respondent) is a categorical data that indicates the income level of respondent before tax and excluding any capital gains or losses [1]. Income is inherently a numeric data, so we will transform these two variables to numeric data ranging from 1 to 7 which represent the income level. Below is a reference table:
```{r, echo=FALSE}
inc_name = c("Less than $25,000","$25,000 to $49,999","$50,000 to $74,999","$75,000 to $99,999","$100,000 to $ 124,999","$125,000 and more")
rank = group_by(gss_numeric, income_respondent) %>% summarise(.groups = 'drop')
rank = rank$income_respondent
tibble(`Income Level`=inc_name, `Numeric Level`=rank) %>% kable(align = 'lccrr', caption='Income Level Reference')
```


__Self rated health__ (code in raw data: SRH_110, name in 'gss.csv': self_rated_health) is a categorical data that indicates what the respondent think his/her health level is [1]. We will also transform this variable to numeric data with range from 1-5. Below is the reference table:
```{r, echo=FALSE}
health_name = c("Poor","Fair","Good","Very good","Execellent")
rank = group_by(gss_numeric, health) %>% summarise(.groups = 'drop')
rank = rank$health
tibble(`Self Rated Health Level`=health_name, `Numeric Level`=rank) %>% kable(align = 'lccrr', caption='Health Level Reference')
```


__Average working hours__ (code in raw data: UHW_16GR, name in 'gss.csv': average_hours_worked) is a categorical data that indicates how many hours the respondent works a week on average [1]. We will convert the variable to numeric with range from 1-5. Below is the reference table:
```{r, echo=FALSE}
hrs_name = c("0 hour","0.1 to 29.9 hours","30.0 to 40.0 hours","40.1 to 50.0 hours","50.1 hours and more")
rank = group_by(gss_numeric, work_hours) %>% summarise(.groups = 'drop')
rank = rank$work_hours
tibble(`Weekly Work Hour Level`=hrs_name, `Numeric Level`=rank) %>% kable(align = 'lccrr', caption='Work Hour Level Reference')

```

### Potential Drawback

For the health variable, we use the self rated health from GSS, but this does not necessarily represent the actual health condition of respondent. The respondent himself/herself might not be a physician and could have wrong understanding of his/her actual health condition.


### Why do we choose these variables?

## Model



We will use a Bayesian linear model.

In other words, we are trying to find a linear formula that describes the relationship between our explanatory variables and the response variable:

$Y_{feeling}=\beta_0+\beta_1\cdot I_{divorced}+\beta_2\cdot I_{CommonLaw} + \beta_3\cdot I_{married} + \beta_4\cdot I_{separated} + \beta_5\cdot I_{widowed} + \beta_6\cdot X_{education} + \beta_7\cdot X_{health} + \beta_8\cdot X_{income} + \beta_9\cdot X_{hours}$

where, 

$\begin{aligned}&Y_{feeling} \text{ is the model's predicted feeling of life as a whole}\\\\&I_i \text{ represents an indicator variable of marital status that can only take value of 0 or 1}\\&\text{For }I_{divorced} \text{, 1 means the person is divorced and 0 means not.}\\&\text{For }I_{CommonLaw} \text{, 1 means the person is living in common law and 0 means not.}\\&\text{For }I_{married} \text{, 1 means the person is married and 0 means not.}\\&\text{For }I_{separated} \text{, 1 means the person is separated and 0 means not.}\\&\text{For }I_{widowed} \text{, 1 means the person is widowed and 0 means not.}\\\\&X_{education} \text{ is a variable with range from 1 to 7, indicating the education level.}\\&X_{health} \text{ is a variable with range from 1 to 5, indicating the self rated health level.}\\&X_{income} \text{ is a variable with range from 1 to 6, indicating the respondent income level.}\\&X_{hours} \text{ is a variable with range from 1 to 5, indicating the average weekily working hour level.}\\\end{aligned}$

For the model to make sense, if one marital status indicator variable $I_i$ is 1, all other $I_i$ should be 0. The variable $I_{single}$ for Single is missing because we use it as a reference variable. In other words, the coefficient $\beta_1$ of $I_{divorced}$ represents how much the respondent's feeling of life as a whole will change, comparing to being Single, if the respondent is divorced, and the same applies to other $I_i$.

We will find $\beta$s that give the best fitted model.

We will assume that all these variables are independent. The difference between a simple linear regression model and a Bayesian one is that, in Bayesian approach, we see the parameters (i.e. $\beta$s) as random variables instead of fixed constant, and the probability distribution of $\beta$s is called a prior distribution.

$$prior=P(\beta_1,...,\beta_9)$$

The objective of Bayesian is to find the conditional distribution of $\beta$ given the sample data (i.e. data in Table 1 ), called posterior distribution. 

$$posterior=P(\beta_1,...,\beta_9|data)$$

By Bayes' rule, the posterior distribution is proportional to the product of likelihood, which is the joint density distribution of all explanatory variables, and prior.

$$posterior\propto prior\cdot likelihood=P(\beta_1,...,\beta_9)\cdot P(data|\beta_1,...,\beta_9)$$

To calculate likelihood, we will just multiply the density distribution of all variables since we assume all variables are independent.

We decide to choose Bayesian because it gives a more compact confidence/credible interval than the simple linear model does [2]. With the same level of confidence, if the interval is smaller, we could give a more precise estimate of the parameter.

We will use the technique of Gibbs Sampling to estimate our coefficients. Since we have multiple coefficients, the distribution of one $\beta$ will condition on the other $\beta$s, and Gibbs sampling provides an iterative algorithm that helps us to estimate each coefficient [3].

```{r, echo=FALSE, include=FALSE}
model <- brm(formula = feelings_life ~ marriage + education + health + income_respondent + work_hours, data = gss_numeric, seed = 84)
```

Population-Level Effects: 
                         Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept                    5.73      0.08     5.58     5.88 1.00     6905     3241
marriageDivorced             0.16      0.05     0.06     0.26 1.00     4634     3340
marriageLivingcommonMlaw     0.52      0.04     0.43     0.60 1.00     4635     3196
marriageMarried              0.73      0.03     0.67     0.79 1.00     4095     3473
marriageSeparated           -0.10      0.07    -0.25     0.04 1.00     5385     2927
marriageWidowed              0.01      0.09    -0.16     0.18 1.00     6403     3262
education                   -0.06      0.01    -0.07    -0.04 1.00     7646     3334
health                       0.54      0.01     0.52     0.57 1.00     8710     3065
income_respondent            0.04      0.01     0.02     0.06 1.00     6006     2952
work_hours                   0.02      0.02    -0.01     0.05 1.00     6828     3298

Family Specific Parameters: 
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     1.39      0.01     1.37     1.40 1.00     9193     2954

Samples were drawn using sampling(NUTS). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).

(Introduce the selected model here. It is expected that you will use some mathematical notation here. If you do please ensure that all notation is explained. You may also want to discuss any special (hypothetical) cases of your model here, as well as any caveats.)

### Marital Status


### Education


### Health


### Income


### Weekily work hours




## Results

Below is a histogram that gives a brief overview of Canadians' feelings of life as a whole in 2017:

```{r, echo=FALSE}
hist(gss_numeric$feelings_life, xlab = 'Feelings of life', main = 'Figure 1')

```

The distribution in Figure 1 is left skewed with a mean of 8.13, so we can say that the overall well-being of Canada in 2017 is quite good.


```{r, echo=FALSE}
ggplot(gss_numeric, aes(x=education, group=education, y=feelings_life)) + geom_boxplot() + ylab("Feelings of Life") + xlab("Education Level") + ggtitle('Figure 2') + theme(plot.title = element_text(hjust = 0.5))
```

Above is a boxplot of education level versus feeling of life. It is not obvious to see whether there is a relationship; we will need the linear model to check the correlations.


```{r}
ggplot(gss_numeric, aes(x=health, group=health, y=feelings_life)) + geom_boxplot() + ylab("Feelings of Life") + xlab("Self Rated Health Level") + ggtitle('Figure 3') + theme(plot.title = element_text(hjust = 0.5))
```

Above is a boxplot of self rated level versus feeling of life. We can see that there is positive correlation between self rated health and feeling of life, but we still need the model to verify this.


```{r}
ggplot(gss_numeric, aes(x=income_respondent, group=income_respondent, y=feelings_life)) + geom_boxplot() + ylab("Feelings of Life") + xlab("Individual Income Level") + ggtitle('Figure 4') + theme(plot.title = element_text(hjust = 0.5))
```

Above is a boxplot of respondent income versus feeling of life. Just by looking at the graph, we cannot see a relationship between these 2 variables.


```{r}
ggplot(gss_numeric, aes(x=work_hours, group=work_hours, y=feelings_life)) + geom_boxplot() + ylab("Feelings of Life") + xlab("Average Weekly Working Hour Level") + ggtitle('Figure 5') + theme(plot.title = element_text(hjust = 0.5))
```

Above is a boxplot of weekly average working hours versus feeling of life. Just through the plot, we might propose that longer work hours is correlated with more satisfied feeling of life; we will need the model to verify this.


```{r}
ggplot(gss_numeric, aes(x=marriage, group=marriage, y=feelings_life)) + geom_boxplot() + ylab("Feelings of Life") + xlab("Marital Status") + ggtitle('Figure 6') + theme(plot.title = element_text(hjust = 0.5))
```

Above is a boxplot of marital status versus feeling of life. Just by looking at the graph, we cannot clearly see a relationship between these 2 variables.




```{r, echo=FALSE}
model
```




(Here you will include all results. This includes descriptive statistics, graphs, figures, tables, and model results. Please ensure that everything is well formatted and in a report style. You must also provide an explanation of the results in this section. You can overflow to an Appendix if needed. 

Please ensure that everything is well labelled. So if you have multiple histograms and plots, calling them Figure 1, 2, 3, etc. and referencing them as Figure 1, Figure 2, etc. in your report will be expected. The reader should not get lost in a sea of information. Make sure to have the results be clean, well formatted and digestible.)




 From the results of Bayesian linear regression, firstly we could conclude if there is any correlation between response variable and each independent variable. The null hypothesis here is that there is no relationship between feelings about life and each factor. If the value of the null hypothesis is in the 95% confidence interval, then the we will not have evidence to reject the null hypothesis; when the confidence interval does not contain the null hypothesis, we could reject the null hypothesis. In other words, the 95% confidence interval for marriage status which are separated and windowed, and for work hours do contain the value of the null hypothesis which is 0. Hence we could conclude that the feelings about life is not correlated with each of these three variables. Secondly, now we could analyze the effects of predictor variables on the outcome variable. It is clear that comparing to the participants who are single, those participantse who got divorced, living common-law and married have better feelings about life, with 0.16 units, 0.52 units, 0.73 units increment respectively. On the contrary, the level of education has a negative impact on the feelings about life because when the educational level increases from lower level to nearest higher level(e.g. the changes from "Less than high school diploma or its equivalent" to "High school diploma or a high school equivalency certificate"), the feelings about life decreases by 0.06 units. Moreover, respondents are 0.54 units more satisfied with life as their health codition moves forward to the nearest better level(such as "good" to "very good"), and the rise from lower income range to the nearest higher income range grows their life satisfaction by 0.04 units.


## Discussion

(Here you will discuss conclusions drawn from the results and comment on how it relates to the original goal of the study (which was specified in the Introduction).)

As we can see from our model, a good health condition greatly promotes personal well-beings in Canada, and it has greater positive impact on well-beings than the individual income has. This result implies that in order to increase the happiness of the society, the government should focus more on public health policies and personal medical support, etc.

Additionally, from the model, we find out that getting married and living common-law also contributes largely to Canadian's feeling of life. This is intuitive since we expect people to be happier when they are in a relationship. This result may imply that the government should improve the common law or simplify the process of marriage registration.

We also discover from the model that higher education level has a tiny negative impact on people's feeling of life. A possible explanation is that as people obtain more knowledge, they might worried about more things. For example, a construction worker may feel very satisfied with his/her job while an economic professor is constantly worrying about an upcoming recession. This does not necessarily mean that government should discourage people from increasing their education level. In fact, a higher education level might relate to a higher income, and by our model, income has a positive influence on the well-being.

One interesting discovery is that, comparing to single respondents, divorced people have a more positive feeling of life. Intuitively, we often relate divorce to grief and distress, but the data from GSS tells a different story. Based on this result, the government may need to change its view on divorce, and consider to modify terms in the marriage laws to address this social issue.

# Weaknesses

When we fit our model, we assume that all variables are independent, but that is not necessarily true. The education level and income may have relations because a good education background possibly helps people to get a high-waged job. Average work hours could also relate to health since if a person works a lot, his/her health condition could be affected.


# Next Steps

(Here you discuss subsequent work to be done after this report. This can include next steps in terms of statistical analysis (perhaps there is a more efficient algorithm available, or perhaps there is a caveat in the data that would allow for some new technique). Future steps should also be specified in terms of the study setting (eg. including a follow-up survey on something, or a subsequent study that would complement the conclusions of your report).)

We could further investigate why divorce has a positive impact on people's feeling of life, and it may require another linear model to discover potential reasons.


## References

[1] Statistics Canada. 2017. General Social Survey on Family (Cycle 31).

[2] Educational Reseatch Techniques. 2018. Linear Regression vs Bayesian Regression. Retrieved October 18, 2020 from https://educationalresearchtechniques.com/2017/10/18/linear-regression-vs-bayesian-regression/

[3] Ilker Yildirim. 2012. Bayesian Inference: Gibbs Sampling. Retrieved October 18, 2020 from http://www.mit.edu/~ilkery/papers/GibbsSampling.pdf

